---
title: "Data Set Generator: Part 1"
description: |
  This posts deals with creating the initial mass fractions to be used as initial
  conditions for the source terms ODE's. 
author:
  - name: Marlon Cabrera-Ormaza
    url: 
date: 06-16-2020
output:
  distill::distill_article:
    self_contained: false
bibliography: ../References/Neural_Networks.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## __Random Initial Mass Fraction Generator__

For the training, testing and validation of our ANN we need to generate a proper data set in such a way that the network can perform supervised learning from it, while keeping grounded on the physics of reality.

In lay terms, this means that it would not make sense to train a network to predict values that are not possible in real life. The intended way to generate the data samples is to solve the system of stiff ordinary differential equations and use the evolution steps of the numerical solver as _input_ and _label_ sets. For this we need to provide first a set of random initial conditions (in this case initial mass fractions and temperature) to pass to an ODE solver. 

It is important that these randomly generated initial conditions agree with was is physically possible. Mathematically speaking, it is possible to generate an infinite number of initial mass fractions containing all the species involved in a given reaction mechanism, however, the existence of some of these might not be actually plausible since they could not agree with one or more conservation law. In particular we'd like our samples to comply with mass conservation.

One way this is feasible is by using Bilger's Mixture Fraction Formula[@Bilger1976]. It is based on the idea that under the assumptions of equal diffusivities for all species, elemental mass fractions are conserved even as the species containing them are being produced or consumed [@KVersteeg2006]:

$$Z_{j}=\sum_{1o=1}^{N} \mu_{i,j} Y_{i}$$
where:

$Z_{j}$ is the conserve mass fraction of element $j$
$\mu_{i,j}$ is the mass fraction of element $j$ in species $i$
$Y_{i}$ is the mass fraction of species $i$
$N$ is the number of species present in the reaction

Then, according to Bilger, for the mixing and combustion of two streams A and B (e.g. fuel and oxidizer), related by a mixture fraction $\xi$ it holds:

$$Z_{j}=Z_{j,A}\xi+(1-\xi)Z_{j,B}$$
where:
$Z_{j,A}$ and $Z_{j,B}$ are elemental mass fractions of element $j$ present in the streams $A$ and $B$.

$\xi$ is the stream mixture fraction

The generated mass fractions must also comply with the condition that their sum must be equal to the unit:

$$\sum_{i=1}^{N}Y_{i}=1$$

## __Setting the work environment__
The libraries that will be used for this are:

- Cantera
- Numpy

```{r}
library(reticulate)
library(knitr)
use_condaenv('tensor')
```


```{python echo=FALSE}
import numpy as np
import cantera as ct

#Set printing format for numpy arrays for ease of reading
np.set_printoptions(formatter={'float': lambda x: "{0:0.4f}".format(x)})
```

## __Analysis of the combustion gas__

Before proceeding to generating the set of initial random condition we need to be able to obtain the chemical properties of our gas mixture. The library _Cantera_ is useful in this endeavor since it allows for the reading of information from a reaction mechanism and then provides functionality for interacting with them via Python.

```{python echo=FALSE}
# We read a solution mechanism and assign its properties to an object
gas = ct.Solution('./Mechanisms/WD_Laubscher.cti')
```

After assigning the properties of a reaction mechanism to an object, _Cantera_ provides functions to use them.

```{python echo=TRUE}
#Examples of some of the Cantera functions to be used:


elements=gas.element_names #Elements present in the mechanism
species=gas.species_names #Species present in the mechanism
print(' Elements in the mechanism:',elements,'\n',
     'Species in the mechanism:',species,'\n',
     'Number of atoms of Oxygen in Methane:'
      ,gas.n_atoms('CH4','O'),'\n',#Number of atoms of an element in a given species
     'Number of atoms of Oxygen in Water:'
      ,gas.n_atoms('H2O','O'),'\n',
      'Mass of Hydrogen: ', #Atomic weight of a given element present in the mechanism
      gas.atomic_weight('H'),'\n',
      'Mass of Methane: ', #Molecular weight of a given species present in the mechanism
      gas['CH4'].molecular_weights[0]
     )
```

Since noble gases do not interact in any reaction (footnote), the following function is designed to remove them:

```{python echo=TRUE}
#Noble gases filter:
def noble_filter(elements):
    filter_elements=[]
    # List of elements to be removed
    noble=['Ar','Kr','Ne','He','Rn','Xe']
    for x in elements:
        if (noble.count(x)==0):
            filter_elements.append(x)
            
    return filter_elements
```

```{python echo=TRUE}
elements=noble_filter(gas.element_names)
print('Elements without noble gases: ',elements)
```


## __Bilger's Definition__

As was detailed above, Bilger's Mixture Fraction Formula provides a nice framework under which initial mas fractions can be created. Provided a random mixture fraction, and elemental mass fraction for two streams, Bilger's Equations could then be solved returning a vector of initial conditions as a result. To illustrate this properly, it is possible to take the previous mechanism (with seven species and four elements)as an example and develop the related system of equations for it. 

The extant species in the mechanism are: $Y_{CH_{4}},Y_{N_{2}},Y_{CO},Y_{CO_{2}},Y_{H_{2}O},Y_{O_{2}},Y_{H_{2}}$; with the elements: $C,H,O,N$ 

Thus we get the following equations: 

$$Y_{CH_{4}}+Y_{N_{2}}+Y_{CO}+Y_{CO_{2}}+Y_{H_{2}O}+Y_{O_{2}}+Y_{H_{2}}=1$$
$$Z_{H}=\mu_{H,H_{2}}Y_{H_{2}}+\mu_{H,H_{2}O}Y_{H_{2}O}+\mu_{H,CH_{4}}Y_{CH_{4}}$$
$$Z_{C}=\mu_{C,CO_{2}}Y_{CO_{2}}+\mu_{C,CO}Y_{CO}+\mu_{C,CH_{4}}Y_{CH_{4}}$$
$$Z_{O}=\mu_{H,H_{2}}Y_{H_{2}}+\mu_{H,H_{2}O}Y_{H_{2}O}+\mu_{H,CH_{4}}Y_{CH_{4}}$$
$$Z_{N}=\mu_{N,N_{2}}Y_{N_{2}}$$

Generalizing for a mechanism with $m$ number of elements and $n$ number of species, this can be rewritten  as a linear system using matrix and vector notation:

$$Ax=b$$

\def\A{
\begin{bmatrix}
    1 \cdots 1 \\
    \mu_{1,1} \cdots \mu_{1,n} \\
    \mu_{2,1} \cdots \mu_{2,n} \\
    \vdots \\
    \mu_{m,1} \cdots \mu_{m,n} 
\end{bmatrix}}

\def\X{
\begin{bmatrix}
    Y_1 \\
    Y_2\\
    Y_3\\
    \vdots \\
    Y_n
\end{bmatrix}}

\def\b{
\begin{bmatrix}
    1 \\
    z_1 \\
    z_2 \\
    \vdots \\
    z_m
\end{bmatrix}}


\begin{equation}
\A_{m+1,n} * \X_{n,1}=\b_{m+1,1   }
\end{equation}

Using the functions described in the previous section possible to build a new one that generates the matrix needed to solve a linear system of equations for the initial conditions that agrees with Bilger's Definition.

To this end, the coefficients $\mu_{i,j}$ can be computed in the following way:

$$\mu_{i,j}=\frac{a_{i,j}W_{j}}{MW_{i}}$$
where:

$a_{i,j}$ is the number of atoms of element $j$ in species $i$ 
$W_{j}$ is the atomic weight of element $j$
$MW_{i}$ is the molecular mass of species $i$


```{python echo=TRUE}
#Matrix for the linear system of Equations:
def bilger_matrix(gas):
    
    #Elements present in the mechanism
    elements=noble_filter(gas.element_names)
    #Species present in the mechanism
    species=gas.species_names
    
    #Matrix to store coefficients
    A=np.ones((len(elements)+1,len(species)))
   
    row=1;
    for x in elements:
        col=0
        for y in species:
            A[row,col]=gas.n_atoms(y,x)*gas.atomic_weight(x)/gas[y].molecular_weights[0]
            col +=1
        row +=1
    return np.copy(A)

A=bilger_matrix(gas)
print(A)
```

As for the resulting vector, it will be created using random mixture fractions of a fuel and an oxidizer. Python dictionaries bestow a simple way to storing this information. For the results to be reproducible, it is necessary to use a seeded random number generator. Nonetheless the implementation of it won't come until later.

```{python echo=TRUE}
#Dictionaries for oxydizer and fuel mass fractions:
fuel={'H':0.03922422,'O':0.19663877,'C':0.11684286,'N':0.64729}
oxy={'H':1.03625e-5,'O':0.23298543,'C':3.194117e-5,'N':0.766970811}

def bilger_vector(f_mix, fuel_dict, oxy_dict, elements):
    
    b=np.ones(len(elements)+1)
    row=1
    for x in elements:
        b[row]=f_mix*fuel_dict[x]+(1-f_mix)*oxy_dict[x]
        row+=1
    return np.copy(b)
```

```{python echo=TRUE}
np.random.seed(0)
rn=np.random.randn(1)
b=bilger_vector(rn,fuel, oxy, elements)
print(b)
```

Now is possible to see both the coefficient matrix and result vector needed to solve the linear system Ax=b. 

```{python echo=TRUE}
print("A =\n\n",A)
print("\nB.transpose =\n\n",b)
```

## __Solving the linear system__

Once the system of equations is set up it is necessary to solve it for the species mass fractions. Unfortunately, for most mechanisms, this will be an undetermined system (i.e. it has more unknowns than equations) and thus, in the case it can be solved, this solution will not be unique. To deal with this issue there are two possibilities using either a [least-squares](https://numpy.org/doc/stable/reference/generated/numpy.linalg.lstsq.html) or or a [non-negative least squares](https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.nnls.html) solver. Both will provide an approximation of the solution by minimizing the squared euclidean norm:

$$\underset{x} {\arg\min}||Ax-b||^2=0$$
whilst  non-negative least squares also implements the restriction:

$$x_i\geq0\quad \forall x_{i} \in x $$

However, they differ in certain aspects, and thus is necessary to analyze the distribution of the samples generated with each in order to select an appropriate method. 

```{python echo=TRUE}
#Solution of the linear system via least squares solver
import pandas as pd
S=[]
F=[]
A=bilger_matrix(gas)
np.random.seed(0)
f_max=1
f_min=0
for i in range(10000):
    f=np.random.random(1)*(f_max-f_min)+f_min
    b=bilger_vector(f[0],fuel, oxy, noble_filter(elements))
    z=np.linalg.lstsq(A, b,rcond=None)[0:]
    S.append(z[0])
    F.append(f[0])
df1=pd.DataFrame(S,columns=gas.species_names)
df1['Mixture']=F
df1.head()
```

```{python echo=TRUE}
#Solution of the linear system via non-negative least squares solver
import pandas as pd
from scipy.optimize import nnls 
S=[]
F=[]
A=bilger_matrix(gas)
np.random.seed(0)
f_max=1
f_min=0
for i in range(10000):
    f=np.random.random(1)*(f_max-f_min)+f_min
    b=bilger_vector(f[0],fuel, oxy, noble_filter(elements))
    z,rnorm=nnls(A,b)
    S.append(z)
    F.append(f[0])
df2=pd.DataFrame(S,columns=gas.species_names)
df2['Mixture']=F
df2.head()
```

### __NNLS vs Least Squares__

In this section, the resulting initial mass fractions will be compared. Using the same mixture fractions, samples are generated with each of the methods and then plotted using a scatter graph. This will allow for a description of their behavior under diferent situations. Things that must be checked are whether any of the methods are producing negative or zero mass fractions for certain species, and if that'd be the case, what is the difference using the other solver for the same random mixture fraction.

#### __CH4__

```{python echo=TRUE}
import json
import plotly
import plotly.express as px
specie=[df1['CH4'],df2['CH4']]
fig = px.scatter(x=specie[0], y=specie[1])
fig.update_layout(
    title="Relation of CH4 Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```
```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
library(plotly)
library(jsonlite)
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```


Here it can be seen that for certain mixture fractions the least squares method is producing negative methane mass fractions which is something that is not desired. Its non negative version produces values in the range between  an 0.05 for the same mixtrue fractions.
#### __O2__



```{python echo=TRUE}
specie=[df1['O2'],df2['O2']]
fig = px.scatter(x=specie[0], y=specie[1])
fig.update_layout(
    title="Relation of O2 Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

None of the solver result in any sort of issue when solving for the $O_{2}$ mass fractions, as both produce only positive values.

### __H2O__


```{python echo=TRUE}
specie=[df1['H2O']]
fig = px.scatter(x=specie[0], y=df2['H2O'])
fig.update_layout(
    title="Relation of H2O Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```
```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

The least squares method produces only positive values, which is the desired behavior. The non-negative solver always results in zero values.

### __N2__

```{python echo=TRUE}
specie=[df1['N2'],df2['N2']]
fig = px.scatter(x=specie[0], y=specie[1])
fig.update_layout(
    title="Relation of N2 Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

Since the mass fraction of $N_2$ does not depend on the other species (and thus only depends in the positive random mixture fraction) it was expected that only non-negative values will be obtained. Nevertheless, the plot above is useful to describe the range in which said variable lies.

### __CO__


```{python echo=TRUE}
specie=[df1['CO'],df2['CO']]
fig = px.scatter(x=specie[0], y=specie[1])
fig.update_layout(
    title="Relation of CO Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

Both methods give only positive results, however the range of NNLS results is more limited, and as can be seen are orders of magnitude smaller.

### __CO2__


```{python}
specie=[df1['CO2']]
fig = px.scatter(x=specie[0], y=df2['CO2'])
fig.update_layout(
    title="Relation of CO2 Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```

```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}

plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

Again similar behavior as in the case for $H_{2}O$. NNLS result only on zero values.

### __H2__


```{python}
import plotly.express as px
specie=[df1['H2'],df2['H2']]
fig = px.scatter(x=specie[0], y=specie[1])
fig.update_layout(
    title="Relation of H2 Mass Fractions obtained by different solvers",
    xaxis_title="Least Squares",
    yaxis_title="Non-negative Least Squares"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

The solutions for $H_2$ act in an alike manner to those for $CO$, where both solvers produce only positive result, but NNLS range is more limited and orders of magnitude smaller.

## __Final Solution of the linear system:__

Since it was seen that the least squares method generates a lot of negative mass fractions, while the non-negative least squares produces a lot of zeroes an obvious idea would be to combine both methods in such a way that they compliment each other. From observing the scatter graphs of the extant species, one could observe that in most cases, whenever a negative value is created using the least squares solver, the non-negative one returns a zero value. Combined in this way the negative mass fractions are dealt with while avoiding that for certain species, only zero fractions are returned (which is the main problem when using nnls).

Using this approach, it is now possible to define a functions that utilizes both solvers.

```{python}
# Define a function to Generate random temperatures and mass fractions

def TY_Generator(seed,samples,gas,fuel_dict,oxy_dict,f_max=0.6,f_min=0.1,T_max=3000,T_min=300):
    S=[]
    F=[]
    A=bilger_matrix(gas)
    np.random.seed(seed)
    T=np.random.random(samples)*(T_max-T_min)+T_min
    np.random.seed(seed)
    for i in range(samples):
        f=np.random.random(1)*(f_max-f_min)+f_min
        b=bilger_vector(f[0],fuel_dict, oxy_dict, noble_filter(gas.element_names))
        z=np.linalg.lstsq(A, b,rcond=None)[0:]
        if (z[0][0]>=0):
            S.append(z[0])
        else:
            z,rnorm=nnls(A,b)
            S.append(z)
    YInit=np.array(S)
    return (np.copy(T),np.copy(YInit))
```

```{python}
# Testing

#Dictionaries for oxydizer and fuel mass fractions:
fuel={'H':0.03922422,'O':0.19663877,'C':0.11684286,'N':0.64729}
oxy={'H':1.03625e-5,'O':0.23298543,'C':3.194117e-5,'N':0.766970811}
#Parameters:
samples=200
gas=ct.Solution('./Mechanisms/WD_Laubscher.cti')
seed=0

T,YInit=TY_Generator(seed,samples,gas,fuel,oxy,f_max=1,f_min=0)
Fractions=pd.DataFrame(YInit,columns=gas.species_names)
```

```{python}
Fractions.head()
```

### __Distribution of the Mass Fractions Samples__

To finish, it is important once again to check the distribution of the mass fractions for each indivifual species.

#### __CH4__:

```{python}
import plotly.figure_factory as ff
from plotly.subplots import make_subplots
specie=[Fractions['CH4']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of CH4 Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

#### __O2__:

```{python}
specie=[Fractions['O2']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of O2 Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

#### __H2O__:

```{python}
specie=[Fractions['H2O']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of H2O Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

#### __N2__:

```{python}
specie=[Fractions['N2']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of N2 Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

#### __CO__:

```{python}
specie=[Fractions['CO']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of CO Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

#### __CO2__:

```{python}
specie=[Fractions['CO2']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of CO2 Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```

#### __H2__:

```{python}
specie=[Fractions['H2']]
group_labels = ['Generator Function']
fig = ff.create_distplot(specie, group_labels,bin_size=.03)
fig.update_layout(
    title="Distribution of H2 Mass Fractions"
);
```

```{python include = FALSE}
fig = json.dumps(fig, cls=plotly.utils.PlotlyJSONEncoder)
```


```{r layout="l-body-outset",echo = FALSE , message=FALSE, warning=FALSE}
plotly::as_widget(
  jsonlite::fromJSON(
    reticulate::py$fig, simplifyVector=FALSE))
```